---
output:
  html_document:
    keep_md: yes
---
Reproducible Research
=====================




Assessment 1 - Introduction
----------------------------


It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.


####Loading the data

```{r load-data}

dataFrame <- read.csv("./data/activity.csv", header = TRUE)

```



####What is mean total number of steps taken per day?

#####For this part of the assignment, you can ignore the missing values in the dataset.

*Calculate the total number of steps taken per day*

```{r tot-steps}

stepsPerDay <- aggregate(steps ~ date, data = dataFrame, sum, na.rm=TRUE)

```

*Make a histogram of the total number of steps taken each day*

```{r hist-plot}

hist(stepsPerDay$steps, main="Total number of steps per day", xlab="Number of steps", ylab= "Frequency in minutes", plot=TRUE)

```

*Calculate and report the mean and median of the total number of steps taken per day*

```{r mean-median-nofilled}
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```

###What is the average daily activity pattern?


*Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)*
```{r time-plot}
DataInterval <- aggregate(steps ~ interval, data = dataFrame, mean, na.rm=TRUE)
with(DataInterval, plot(interval, steps,type="l", ylab= "Average Steps x Day", xlab="5 min Interval", col="blue"))
title(main="Activity Pattern")
```

*Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?*
```{r max-interval}
DataInterval[which.max(DataInterval$steps), ]$interval
```

###Inputing missing values

Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data

*Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)*

```{r tot-missing}
sum(is.na(dataFrame$steps))
```

*Devise a strategy for filling in all of the missing values in the dataset.*

My strategy  will be based, on the to copy of the results from the mean calculations obtained on the "Data Interval" data frame.
Each value on the same interval will be assigned to the NA value on the new DF.


*Create a new dataset that is equal to the original dataset but with the missing data filled in.*
```{r filled-dataFrame}
# Make a new dataset with the original data
dataFrameFilled <- dataFrame
count = 0 
for (i in 1:nrow(dataFrameFilled))
{
  # Check for NA values
  if (is.na(dataFrameFilled[i, ]$steps))
  {
    # Assign the step value for the corresponding interval
    dataFrameFilled[i, ]$steps <- DataInterval[DataInterval$interval== dataFrameFilled[i, ]$interval, ]$steps
    count = count + 1
  }
}

```

*Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.*

```{r hist-plot-filled}
stepsPerDay2 <- aggregate(steps ~ date, data = dataFrameFilled, sum, na.rm=TRUE)
hist(stepsPerDay2$steps, main="Total number of steps per day", xlab="Number of steps", ylab= "Frequency in minutes", plot=TRUE)
```

*Calculate the mean and median*

```{r mean-median-filled}
mean(stepsPerDay2$steps)
median(stepsPerDay2$steps)
```

**Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?**

The mean value is the same as the value before imputing the missing values to NA. The median shows a minor difference.

###Are there differences in activity patterns between weekdays and weekends?

*Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.*

```{r new-column-dataframe}
dataFrameFilled$dayType <- ifelse(weekdays(as.Date(dataFrameFilled$date,format="%Y-%m-%d")) %in% c("Satuday", "Sunday"), 
                              "weekend", "weekday")
```

*Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).*

```{r panel-plot}
library(lattice)
dataFrameWeekday = aggregate(steps ~ interval + dayType, dataFrameFilled, mean)
xyplot(steps ~ interval | factor(dayType), data = dataFrameWeekday, aspect = 1/2, 
       type = "l")
```